---
title: "join"
---


```{r setup}
library(tidyverse)
wd_events=readRDS("data/wd_events.RDS")
wd_full=readRDS("data/wd_full.RDS")
```

# Wikidata Map

```{r wd_map, warning=FALSE, message=FALSE}
library(leaflet) 
 wd_map=wd_full %>%
  # filter NA coords before transforming in sf multipoint
  filter(!is.na(coords)) %>%
  sf::st_as_sf(wkt="coords") %>% 
  # group by flood event
  group_by(flood,flood_label) %>%
  # in case there is no Wikipedia article 
  mutate(noarticle=all(is.na(article)),
         noimage=all(is.na(image)),
         lang=stringr::str_extract(article,"..(?=\\.wikipedia)"),
         flood=stringr::str_replace(flood,"wd:","")) %>% 
  mutate(article=case_when(noarticle~"",
                           !noarticle~glue::glue("<a href='{article}',target='_blank'>{lang} 🔗 </a>")),
         image=case_when(noimage~"",
                         !noimage~glue::glue("<img src='{image}'  width='200'>"))) %>% 
  summarise(coords_from=first(coords_from),
            year=first(year),
            country_label=first(country_label),
            deathtoll=mean(deathtoll,na.rm=TRUE),
            date=first(date),
            start=first(start),
            end=first(end),
            date_precision=first(date_precision),
            article=paste0(unique(article), collapse=" "),
            image=paste0(unique(image),collapse=" ")) %>% 
  ungroup() %>% 
  mutate(popup=glue::glue("<h1>{flood_label}<a href='http://www.wikidata.org/entity/{flood}'
                             target='_blank'>🔗</a></h1>")) %>%
  mutate(popup=case_when(!is.na(date)~glue::glue("{popup}<p>date: {date}</p>"),
                                 TRUE~popup)) %>% 
  mutate(popup=case_when(!is.na(deathtoll)~glue::glue("{popup}<p>deathtoll:{deathtoll}</p>"),
                                 TRUE~popup)) %>% 
  mutate(popup=case_when(!is.na(article)~glue::glue("{popup}<p>{article}</p>"),
                                 TRUE~popup)) %>% 
  mutate(popup=case_when(!is.na(image)~glue::glue("{popup}<p>{image}</p>"),
                                 TRUE~popup)) %>% 
  sf::st_centroid()
coords=wd_map %>%
  sf::st_coordinates() %>% 
  as_tibble() %>% 
  select(long=X,lat=Y)
wd_map=wd_map %>% 
  bind_cols(coords) %>% 
  mutate(coords_txt=as.character(coords))
```

```{r def_jitter_coords}
jitter_coord=function(datlonlat,n){
  long=datlonlat$long+runif(n,-1,1)
  lat=datlonlat$lat+runif(n,-1,1)
  result=tibble::tibble(flood=datlonlat$flood,
                           long=long,
                           lat=lat)
  return(result)
}
```

```{r jittered_coords}
jittered_coords=wd_map  %>% 
  sf::st_drop_geometry() %>% 
  select(flood,coords_txt,long,lat) %>% 
  unique() %>% 
  group_by(coords_txt) %>%
  mutate(n=n()) %>%
  tidyr::nest(data=c(flood,long,lat)) %>%
  mutate(data=purrr::map2(data,n,jitter_coord)) %>%
  tidyr::unnest(data) %>%
  ungroup()

```

```{r}
wd_map=wd_map %>%
  sf::st_drop_geometry() %>% 
  select(-long,-lat,-coords_txt) %>% 
  left_join(jittered_coords %>%
              select(flood,long,lat),
            by="flood") %>% 
  sf::st_as_sf(coords = c("long", "lat"), 
               crs = 4326, agr = "constant")
```

```{r}
# Définition d'une échelle colorée 
# (en fonction de date de sortie) 
pal <- colorNumeric(c("red", "green", "blue"),
                    c(1648,1900,1950,1980,2000,2010,2023)) 
# Création de la carte 
leaf_wd_map=leaflet(wd_map) %>% # déf carte 
  addProviderTiles(providers$Esri.NatGeoWorldMap) %>% # ajout fond de carte
  addCircleMarkers(col=~pal(year),
                   popup = ~popup,
                   radius =~log(deathtoll+2)
                   ) 

```

```{r leaf_wd_map, fig.width=10,fig.height=8}
leaf_wd_map 
```



# Data from institutionnally curated databases

## Data from the Dartmouth Flood Observatory

```{r def_avdate}
avdate=function(date1,date2){
  date1=as.numeric(date1)
  date2=as.numeric(date2)
  result=as.Date(mean(c(date1,date2)))
  return(result)
}
```

```{r read_tib_dfo, warning=FALSE, message=FALSE}
dfo_raw <- readr::read_csv("data/FloodArchive.csv", 
    locale = locale(decimal_mark = ","))
colnames(dfo_raw)=c("index","GlideNumber","dfo_country","dfo_other_country",
           "dfo_centroid_x","dfo_centroid_y","area","dfo_began","dfo_ended",
           "validation","dfo_dead","displaced","maincause","severity")
dfo_raw=dfo_raw %>% 
  mutate(start=lubridate::dmy(dfo_began),
         end=lubridate::dmy(dfo_ended)) %>% 
  mutate(dfo_other_country=case_when(is.na(dfo_other_country)~"",
                                     dfo_other_country==0~"",
                                     TRUE~dfo_other_country)) %>% 
  mutate(flood=as.character(index),
         flood_label=as.character(index),
         country_label=paste(dfo_country," ",dfo_other_country),
         deathtoll=dfo_dead,
         coords=paste0("Point(",dfo_centroid_x," ",dfo_centroid_y,")"))
dfo_comp=dfo_raw %>% 
  select(flood,flood_label,country_label,deathtoll, start, end,coords) %>% 
  mutate(date=purrr::map2(start,end,avdate)) %>% 
  tidyr::unnest(date)
```

```{r}
coords_wd_map=sf::st_coordinates(wd_map) %>%
  as_tibble() %>%
  mutate(coords=glue::glue("Point({X} {Y})")) %>%
  pull(coords)
wd_comp=wd_map %>%
  sf::st_drop_geometry() %>% 
  mutate(coords=coords_wd_map) %>% 
  select(flood,flood_label,country_label,deathtoll, date ,start, end,coords,popup) %>% 
  mutate(flood=stringr::str_replace(flood,"wd:",""))
```

This dataset documents `r nrow(dfo_comp)` flood events that occurred between `r min(dfo_comp$start)` and `r max(dfo_comp$end)`.

```{r}
Nwd_in_range=wd_events %>% 
  filter(year>=1985 & year<=2021) %>% 
  sf::st_drop_geometry() %>%
  summarise(n=length(unique(flood))) %>% 
  pull(n)
```

Based on the dates of observations for the DFO data base, `r Nwd_in_range` out of the `r nrow(wd_events)` flood events in our Wikidata base might fall into it.

## Basic stats

```{r}
wd_comp=wd_comp  %>% 
   mutate(source="wd") %>% 
   mutate(flood=paste0("wd:",flood)) %>% 
   sf::st_drop_geometry() %>% 
   filter(date>lubridate::ymd("1985-01-01") & 
          date<lubridate::ymd("2021-12-31")) 
dfo_comp=dfo_comp %>%
   mutate(source="dfo") %>%
   mutate(popup=glue::glue("<h1>{flood_label}</h1>
                            <p>date: {date}</p>
                            <p>deathtoll:{deathtoll}</p>"))  
tib=bind_rows(wd_comp,dfo_comp) %>% 
  mutate(date=lubridate::round_date(date,"month")) %>% 
   sf::st_as_sf(wkt="coords")
```

Based on the spatial and temporal data in both datasets, we try to find correspondences between the events described in Wikidata and those described in the Dartmouth Flood Observatory.

```{r def_find_corresponding_flood}
# tib %>% 
#   sf::st_transform(crs="")

find_corresponding_flood=function(flood_id,tib){
  tib_sub=tib %>% filter(flood==flood_id)
  tib_dist=tib %>% 
    mutate(disttime=abs(tib$date-tib_sub$date),
           distspace=sf::st_distance(tib,tib_sub)[,1]) %>% 
    filter(flood!=flood_id) %>%  
    filter(disttime==min(disttime)|distspace==min(distspace)) %>% 
    filter(disttime<60 & distspace<2) %>%
    select(floodcorr=flood,
           disttime,
           distspace,
           sourcecorr=source,
           country_corr=country_label,
           deathtoll_corr=deathtoll) %>% 
    sf::st_drop_geometry()
  if(nrow(tib_dist)==0){
    tib_dist=tibble::tibble(floodcorr=NA,
                            disttime=NA,
                            distspace=NA,
                            sourcecorr=NA,
                            countrycorr=NA,
                            deathtollcorr=NA)
  }
  return(tib_dist)
}
```

```{r}
if(!file.exists("data/tibcorr.RDS")){
tibcorr= tib %>% 
  mutate(data=purrr::map(flood,find_corresponding_flood,tib=tib)) %>% 
  tidyr::unnest(data) %>% 
  filter(source=="wd" & sourcecorr=="dfo")
saveRDS(tibcorr,
        "data/tibcorr.RDS")
}
```

```{r}
tib=sf::st_as_sf(tib,wkt="coords")
tibcorr=readRDS("data/tibcorr.RDS") %>% 
  sf::st_drop_geometry()
joined_events=bind_rows(tibble::tibble(id=1:nrow(tibcorr),flood=tibcorr$flood),
               tibble::tibble(id=1:nrow(tibcorr),flood=tibcorr$floodcorr))
joined_events=tib %>% 
  left_join(joined_events, by="flood") %>%  
  select(id,flood) %>% 
  group_by(id) %>% 
  summarise(m =mean(id),do_union=FALSE) %>% 
  sf::st_cast("LINESTRING")
joined_events=joined_events[1:(nrow(joined_events)-1),]
```

## Map comparison

```{r}
library(leaflet) 
tib_comp_map=tib %>%
  sf::st_as_sf(wkt="coords")
# Définition d'une échelle colorée 
# (en fonction de date de sortie) 
# Création de la carte 
pal=colorFactor(c("red","blue"), domain=c("wd","dfo"))
comp_map=leaflet(tib_comp_map)  %>%
  addProviderTiles(providers$Esri.NatGeoWorldMap) %>% # ajout fond de carte
  addCircleMarkers(col=~pal(source),
                   popup =~popup,
                   radius =~log(deathtoll+2)
                   )  %>%
  addPolylines(data=joined_events,color="green")
  

```

```{r}
comp_map
```
